### 2022-04-13 paper/Chapters/Evaluation.tex
Prepare Complete (not final) draft
This draft will be sent of to each of the committee members to read through before the thesis defense.
The PDF has been rendered in "final" mode, so no comments will show through.

small.pdf is a compressed version of the main pdf, with no differences except each of the images have been compressed to 300dpi.
Feedback from advisor and thesis defense

\chapter{Evaluation} % Main chapter title

\label{Chapter6} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

@@

\begin{enumerate}
  \item Both the host and client computers are fully rebooted to ensure no other application is running.
  \item The host computer starts up a single browser window with the web page described in Section \ref{sec:DevelopingTestingAndMeasurementTools}.
  \item If the test calls for the CPU or the GPU to be stressed, the application \enquote{Blender} is started on the host machine, and a benchmark is performed for the CPU or the GPU.
        This benchmark will run for longer than the duration of the test, and will keep the CPU or GPU at 100\% load for the entire duration of the benchmark.
  \item Both devices start up their respective data collection tool that will record key-strokes and changes in the screen's color.
  \item The client machine connects to the host computer using the application being tested.
  \item The user presses the \enquote{0} key on the client's numpad to synchronize the two devices' data recording tools.
  \item The user begins by typing on the client's keyboard slowly, allowing both computers to record their data.
  \item The user then types on the client's keyboard quickly to ensure that quick successive inputs are handled correctly and no data is lost or duplicated.
  \item Once typing is complete, both the client and the host computers close their data recording tools and the client ends the streaming session.
\end{enumerate}

@@

      \legend{Ideal,CPU stressed,GPU stressed}
    \end{axis}
  \end{tikzpicture}
  \caption[Color Delay Data]{Time for client to render update to the screen.}
  \label{fig:ColorDelay}
\end{figure}

@@

Apart from the performance of each application, the quality of the user experience must also be taken into account when considering each application.
When analyzing the forked version of Moonlight, specific quantitative data described in Section \ref{sec:DevelopingTestingAndMeasurementTools} can be collected directly from the application while it is running.
Once the streaming session concludes, the statistics are collated into human-readable numbers and presented in the console.
What follows is an example output from one of the test streaming sessions:

@@

\begin{figure}[h]
  % \centering
  % \begin{subfigure}{.25\textwidth}
  %   \centering
  %   \includegraphics[width=1\linewidth]{Figures/crd/bandwidth}
  %   \caption{Bandwidth usage in bits per second}
  %   \label{fig:crdBandwidth}
  % \end{subfigure}%
  \begin{subfigure}{.33\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{Figures/crd/delay}
    \captionsetup{width=.85\linewidth}
    \caption{delay of image rendering in milliseconds}
    \label{fig:crdDelay}
  \end{subfigure}%
  \begin{subfigure}{.33\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{Figures/crd/fps}
    \captionsetup{width=.85\linewidth}
    \caption{Frames per second encoded by host and decoded by client}
    \label{fig:crdFPS}
  \end{subfigure}%
  \begin{subfigure}{.33\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{Figures/crd/quality}
    \captionsetup{width=.85\linewidth}
    \caption{VP8 compression quality of current frames. Lower means CRD is struggling to keep up}
    \label{fig:crdQuality}
  \end{subfigure}
  \captionsetup{width=.8\linewidth}
  \caption{Statistics from CRD during the most demanding test session. Data recorded over 60 seconds.}
  \label{fig:crdStats}
\end{figure}

Unfortunately, as there is no official documentation for this data, analysis on these figures is limited but insights on the stream quality can still be found.
As seen in Figure \ref{fig:crdDelay}, the delay of image rendering climbs as the CPU is put under stress, with frames sometimes taking entire seconds to be rendered on the client machine.
It is unclear on how much of the delay is due to the host machine versus the client machine, as the delay obviously rises as the host's CPU is put under stress, but it is clear that as the host machine is under heavy load the frames can take entire seconds to be rendered on the client machine.
Figure \ref{fig:crdQuality} shows the quality setting of the VP8 compression algorithm being adjusted over time to keep up with demand.
This is a decent indication of the quality of the video stream over time, as the quality drops in an attempt to keep the stream updating the client with as much frequency as possible instead of sending full frames.
With the knowledge that CRD leverages the VP8 compression algorithm \cite{miniorange_chromoting}, it is possible to infer than as the quality goes down, the frames being sent contain less information about the current frame and instead more temporal data to construct the new frame from the previous frame.
This is also seen with visual smudging on the client's screen as the stream slows down.

@@

By combining the information from the previous two sections, a wholistic picture of the performance of each application can be drawn. Firstly, Virtual Network Computing works as a bare-bones method of connecting to a host machine from a client device in the simplest way possible, but struggles at keeping up with even mildly demanding tasks.
As seen in Figure \ref{fig:ColorDelayLoss}, VNC failed to provide a smooth experience for the user even in ideal conditions by simply not rendering updates to the screen in time for the next frame to be shown.
This, combined with a slow response time and delayed inputs, causes a poor experience for the user when attempting to use demanding applications or anything that requires speed.

Secondly, Chrome Remote Desktop performs well under idea conditions and even succeeds at at maintaining steady performance under some GPU load, but it struggles to keep up whenever the host's CPU is under stress.
Due to CRD's reliance on web technologies, it too suffers from the same limitations as seen in other web streaming applications such as visual smearing and dependence on the CPU.
In fact, these issues are something technology companies like Google have been trying so solve, with the release of their new VP9 codec used for YouTube starting in 2014, but such technology has not yet been introduced to CRD \cite{google_io_vp9}.
Unfortunately, due to the dependence on the CPU, and difficulty in handling conditions where the host computer is running strenuous tasks, Chrome Remote Desktop is not a suitable software solution for the purposes of this paper.

@@

Finally, Moonlight proves to be a high-performant solution that isn't limited by the host's CPU or GPU being under stress.
Each test resulted in highly similar results, and while it does on average have slightly higher response times compared to RDP (Figures \ref{fig:InputDelay} and \ref{fig:ColorDelay}), it only falls behind by 5-10ms -- \todoquestion{is this appropriate?}quicker than the blink of an eye.
By not dropping any frames, and numerically proving it's ability to keep the stream above 60 frames per second, Moonlight proves to be a great solution for streaming demanding applications to a remote client without compromising its usability.

>>>>

\chapter{Evaluation}

\label{Chapter6}

\todoadvisor{How many samples were taken to create the figures? How many times was the test run?}

@@

\begin{enumerate}
  \item Both the host and client computers are fully rebooted to ensure no other application is running.
  \item The host computer starts up a single browser window with the web page described in Section \ref{sec:DevelopingTestingAndMeasurementTools}.
  \item If the test calls for the CPU or the GPU to be stressed, the application \enquote{Blender} is started on the host machine, and a benchmark is performed for the CPU or the GPU.
        This benchmark will run for longer than the duration of the test, and will keep the CPU or GPU at 100\% load for the entire duration of the benchmark.
  \item Both devices start up their respective data collection tool that will record key-strokes and changes in the screen's color.
  \item The client machine connects to the host computer using the application being tested.
  \item The user presses the \enquote{0} key on the client's numpad to synchronize \todoadvisor{Syncronize what?}the two devices' data recording tools.\todoadvisor{How far off are they?}
  \item The user begins by typing on the client's keyboard \todoadvisor{what does slowly mean?}slowly, allowing both computers to record their data.
  \item The user then types on the client's keyboard quickly to ensure that \todoadvisor{Why is there a seperation here?}quick successive inputs are handled correctly and no data is lost or duplicated.
  \item Once typing is complete, both the client and the host computers close their data recording tools and the client ends the streaming session.
\end{enumerate}

@@

      \legend{Ideal,CPU stressed,GPU stressed}
    \end{axis}
  \end{tikzpicture}
  \caption[Color Delay Data]{Time for client to render update to the screen.}
  \label{fig:ColorDelay}
\end{figure}
\todoadvisor{Maybe put full figure in the appendix?}

@@

Apart from the performance of each application, the quality of the user experience must also be taken into account when considering each application.
Each application will be assessed individually, and then compared in Section \ref{sec:EvaluationSummary}
When analyzing the forked version of Moonlight, specific quantitative data described in Section \ref{sec:DevelopingTestingAndMeasurementTools} can be collected directly from the application while it is running.
Once the streaming session concludes, the statistics are collated into human-readable numbers and presented in the console.
What follows is an example output from one of the test streaming sessions:

@@

\begin{figure}[h]
  % \centering
  % \begin{subfigure}{.25\textwidth}
  %   \centering
  %   \includegraphics[width=1\linewidth]{Figures/crd/bandwidth}
  %   \caption{Bandwidth usage in bits per second}
  %   \label{fig:crdBandwidth}
  % \end{subfigure}%
  \begin{subfigure}{.33\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{Figures/crd/delay}
    \captionsetup{width=.85\linewidth}
    \caption{delay of image rendering in milliseconds}
    \label{fig:crdDelay}
  \end{subfigure}%
  \begin{subfigure}{.33\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{Figures/crd/fps}
    \captionsetup{width=.85\linewidth}
    \caption{Frames per second encoded by host and decoded by client}
    \label{fig:crdFPS}
  \end{subfigure}%
  \begin{subfigure}{.33\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{Figures/crd/quality}
    \captionsetup{width=.85\linewidth}
    \caption{VP8 compression quality of current frames. Lower means CRD is struggling to keep up}
    \label{fig:crdQuality}
  \end{subfigure}
  \captionsetup{width=.8\linewidth}
  \caption[CRD Data]{Statistics from CRD during the most demanding test session. Data recorded over 60 seconds.}
  \label{fig:crdStats}
\end{figure}

Unfortunately, as there is no official documentation for this data, analysis on these figures is limited but insights on the stream quality can still be found.
As seen in Figure \ref{fig:crdDelay}, the delay of image rendering climbs as the CPU is put under stress, with frames sometimes taking entire seconds to be rendered on the client machine.
It is unclear on how much of the delay is due to the host machine versus the client machine, as the delay obviously rises as the host's CPU is put under stress, but it is clear that as the host machine is under heavy load the frames can take entire seconds to be rendered on the client machine.
Figure \ref{fig:crdQuality} shows the quality setting of the VP8 compression algorithm being adjusted over time to keep up with demand.
This is a decent indication of the quality of the video stream over time, as the quality drops in an attempt to keep the stream updating the client with as much frequency as possible instead of sending full frames.
With the knowledge that CRD leverages the VP8 compression algorithm \cite{miniorange_chromoting}, it is possible to infer that as the quality goes down, the frames being sent contain less information about the current frame and instead more temporal data to construct the new frame from the previous frame.
This is also seen with visual smudging on the client's screen as the stream slows down.

@@

By combining the information from the previous two sections, a wholistic picture of the performance of each application can be drawn. Firstly, Virtual Network Computing works as a bare-bones method of connecting to a host machine from a client device in the simplest way possible, but struggles at keeping up with even mildly demanding tasks.
As seen in Figure \ref{fig:ColorDelayLoss}, VNC failed to provide a smooth experience for the user even in ideal conditions by simply not rendering updates to the screen in time for the next frame to be shown.
This, combined with a slow response time and delayed inputs, causes a poor experience for the user when attempting to use demanding applications or anything that requires speed, making it unsuitable for the purposes of this paper.

Secondly, Chrome Remote Desktop performs well under ideal conditions and even succeeds at at maintaining steady performance under some GPU load, but it struggles to keep up whenever the host's CPU is under stress.
Due to CRD's reliance on web technologies, it too suffers from the same limitations as seen in other web streaming applications such as visual smearing and dependence on the CPU.
In fact, these issues are something technology companies like Google have been trying so solve, with the release of their new VP9 codec used for YouTube starting in 2014, but such technology has not yet been introduced to CRD \cite{google_io_vp9}.
Unfortunately, due to the dependence on the CPU, and difficulty in handling conditions where the host computer is running strenuous tasks, Chrome Remote Desktop is not a suitable software solution for the purposes of this paper.

@@

Finally, Moonlight proves to be a high-performant solution that isn't limited by the host's CPU or GPU being under stress.
Each test resulted in highly similar results, and while it does on average have slightly higher response times compared to RDP (Figures \ref{fig:InputDelay} and \ref{fig:ColorDelay}), it only falls behind by 5-10ms -- quicker than the blink of an eye.
By not dropping any frames, and numerically proving it's ability to keep the stream above 60 frames per second, Moonlight proves to be a great solution for streaming demanding applications to a remote client without compromising its usability.

### 2022-04-13 paper/Chapters/Hardware.tex
Feedback from advisor and thesis defense

This section will explain the different constraints taken into account while developing the hardware portion of this project.
First, the hardware must be performant enough to provide a positive user experience (\ref{subsec:HardwarePerformance}), and secondly the product must be cheap enough to justify it over similar alternatives (\ref{subsec:HardwareCost}).

@@

Finally, everything must connect to the CM4.
Figure \ref{fig:CM4Circuit} shows the the two mezzanine connectors the CM4 uses to communicate with the rest of the PCB (schematic in Figure \ref{fig:pcb_schematic_cm4}).
Remember that there is only 0.04mm of space in between each of the pins, leaving very little space to route the traces where they need to go.
The first thing to notice is the large stretch of traces on the back side of the board coming from the HDMI ports.
Because most of the traces come in positive and negative pairs, called differential pairs, the two tracks must be roughly the same length as each other.
This is because as data is transmitted across the pair of tracks, they need to arrive at their destination at the same time based on the physical distance the signals need to cover.
Since the speed of electricity is constant throughout the PCB board, the traces may need to wind back and forth so that the lengths of the positive and negative tracks are equal.
Also note the blank area surrounded by vias at the top of Figure \ref{fig:CM4Circuit}.
This is where the WiFi and Bluetooth antennas are located on the CM4 board.
By removing the copper pour from this area on the PCB board and stitching the perimeter with vias, it prevents the board from acting like another antenna and interfering with the signals being sent to and from the CM4.
Finally, each of the other components are connected to the CM4 board in the same manner as described throughout this chapter, such as the Ethernet adapter, the SD card reader, and remaining ports.
\todoadvisor{Why are those components not described in the same way as power/USB/etc.?}\todoquestion{The Ethernet port is hooked directly into the CM4 connector with nothing interested except that they are also differential pairs, as described in the HDMI section. I suppose they could be explained fully, but I worry that it wouldn't add any useful information.}
See Appendices \ref{AppendixA} and \ref{AppendixB} for further information.

@@

The second board was assembled mostly successfully, and allowed the first stage of testing to begin.
The only issue apart from some ports being too close together with real-world cables was one component having the wrong footprint on the board.
Unfortunately, the component \emph{U2 (MCP73871-2CCI\_ML)}, is described as using the \emph{QFN-20-1EP\_4x4mm} footprint in it's datasheet\todoquestion{Does this need to be cited? Can you cite a datasheet like this?}, but actually uses a revised version of the footprint, shown in Figure \ref{fig:FootprintQFN204x4}.
This discrepancy meant that that component could not be soldered, and it's associated circuit would not function properly.
In this case, the component was a part of the battery charging circuit, causing the entire primary power supply to be unusable.
Thankfully, as described in Section \ref{sec:Prototyping}, the secondary method of powering the board could be used to test the remaining features of the board.
After making the necessary fixes, a second round of five PCB boards and a stencil was ordered.
Because there was still enough components to assemble three more boards, and no components needed to be changed, no more components were ordered at this stage.
Once the boards were delivered, the revised board was again sent to be professionally assembled.

>>>>

This section will explain the different constraints taken into account while developing the hardware portion of this project.
\todoadvisor{Target user or audience of the product should be described more accurately. Since no studies or research has been conducted regarding which group of people are interested in / will use this product, no specific requirements can be derived for the system.
At this point the only source of the requirements is the author of the thesis and hence it should be worded this way. The thesis should initially state that the requirements for the user experience are based on the authors requirements/preferences, to make clear that no other users were involved or a study conducted with other persons.}
First, the hardware must be performant enough to provide a positive user experience (\ref{subsec:HardwarePerformance}), and secondly the product must be cheap enough to justify it over similar alternatives (\ref{subsec:HardwareCost}).

@@

Finally, everything must connect to the CM4.
Figure \ref{fig:CM4Circuit} shows the the two mezzanine connectors the CM4 uses to communicate with the rest of the PCB (schematic in Figure \ref{fig:pcb_schematic_cm4}).
Remember that there is only 0.04mm of space in between each of the pins, leaving very little space to route the traces where they need to go.
The first thing to notice is the large stretch of traces on the back side of the board coming from the HDMI ports.
Because most of the traces come in positive and negative pairs, called differential pairs, the two tracks must be roughly the same length as each other.
This is because as data is transmitted across the pair of tracks, they need to arrive at their destination at the same time based on the physical distance the signals need to cover.
Since the speed of electricity is constant throughout the PCB board, the traces may need to wind back and forth so that the lengths of the positive and negative tracks are equal.
Also note the blank area surrounded by vias at the top of Figure \ref{fig:CM4Circuit}.
This is where the WiFi and Bluetooth antennas are located on the CM4 board.
By removing the copper pour from this area on the PCB board and stitching the perimeter with vias, it prevents the board from acting like another antenna and interfering with the signals being sent to and from the CM4.
Finally, each of the other components are connected to the CM4 board in the same manner as described throughout this chapter, such as the Ethernet adapter, the SD card reader, and remaining ports.
\todoadvisor{Why are those components not described in the same way as power/USB/etc.?}\todoquestion{The Ethernet port is hooked directly into the CM4 connector with nothing interested except that they are also differential pairs, as described in the HDMI section. I suppose they could be explained fully, but I worry that it wouldn't add any useful information.}\todoadvisor{Explain why it's skimed over}
See Appendices \ref{AppendixA} and \ref{AppendixB} for further information.

@@

The second board was assembled mostly successfully, and allowed the first stage of testing to begin.
The only issue apart from some ports being too close together with real-world cables was one component having the wrong footprint on the board.
Unfortunately, the component \emph{U2 (MCP73871-2CCI\_ML)}, is described as using the \emph{QFN-20-1EP\_4x4mm} footprint in its datasheet\todoquestion{Does this need to be cited? Can you cite a datasheet like this?}\todoadvisor{Yes}, but actually uses a revised version of the footprint, shown in Figure \ref{fig:FootprintQFN204x4}.
This discrepancy meant that that component could not be soldered, and it's associated circuit would not function properly.
In this case, the component was a part of the battery charging circuit, causing the entire primary power supply to be unusable.
Thankfully, as described in Section \ref{sec:Prototyping}, the secondary method of powering the board could be used to test the remaining features of the board.
After making the necessary fixes, a second round of five PCB boards and a stencil was ordered.
Because there was still enough components to assemble three more boards, and no components needed to be changed, no more components were ordered at this stage.
Once the boards were delivered, the revised board was again sent to be professionally assembled.

### 2022-04-13 paper/Chapters/Software.tex
Feedback from advisor and thesis defense

\chapter{Developing the Software} % Main chapter title

\label{Chapter5} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

The second half of the project is composed of software that is used to connect the client machine to a host computer.
Designing the software begins with identifying it's requirements in Section \ref{sec:SoftwareRequirements}.
Then, potential avenues for development are discussed in Section \ref{sec:PotentialAvenues} so that the software does not need to be built from scratch.
Following that, special considerations must be made while developing the software for the CM4 running on an \emph{ARM} processor in Section \ref{sec:DevelopingForARM}, and the development of the suite of tools used to record data is detailed in Section \ref{sec:DevelopingTestingAndMeasurementTools}.

@@

Unfortunately, the Nvidia GameStream protocol is both proprietary and a closed source project.
Very little is officially known about the protocol, other than it uses specific video encoders built into Nvidia Graphics cards from the 600 series from 2012 and newer \cite{gamestream_userguide}.
Apart from that, everything has been researched and reverse engineered by the community.
A GameStream session is started by first connecting to a host computer which has enabled the GameStream service through Nvidia's GeForce Experience application.
Once the service is enabled, the computer will listen on port \emph{47989}\todoquestion{Should this be italicized? quoted? plaintext?} for the following web requests:
\begin{itemize}
  \item /serverinfo
  \item /pair
  \item /unpair
  \item /applist
  \item /launch
  \item /resume
  \item /cancel
\end{itemize}
Once a session has been started with a \emph{/launch} request, raw data covering video, audio, and input is passed back and forth over a number of ports until the session is terminated.
An end-to-end example of the protocol's usage is as follows:

@@

Moonlight is an unofficial third-party open-source client for the Nvidia GameStream protocol \cite{moonlight}.
The project is structured as a core library written in \emph{C} with a number of community built clients built on top of it for various platforms.
In essence, it turns the process described in the previous section (\ref{subsec:NVIDIAGameStream}) into a usable programming interface.
By reverse engineering the protocol originally implemented by the Nvidia Shield, Moonlight enables the development of a GameStream client for almost any operating system.
The Moonlight community has already developed fantastic clients for many platforms such as Windows, Mac, Linux, Android, iOS, Web, and even other gaming consoles.
All that is needed for the purposes of this project is to make a fork of the project that tweaks it to best suit the CM4 that drives the hardware and develop a suite of tools to test it's functionality.

@@

In order to create an out-of-the-box experience for someone using the hardware for this project, a custom operating system image is created that has all the dependencies, libraries, and applications pre-installed and pre-configured.
Manjaro provides good tools for building and managing custom images of the Manjaro operating system for \emph{ARM} architectures \cite{manjaro_arm_tools}, with the only drawback being that it is that it must be run on an \emph{ARM} processor, and is much more difficult to run on another architecture.
To make development easier, a toolchain was built that could be run on the CM4 to build the code that is written on other computers.
All the developer needs to do is to pull the git repository of the project, run a set up script to install dependencies, and then run the build script to compile everything into an immediately deployable OS image.
This build script will automatically compile all the necessary code using \emph{cmake}, package all the necessary files using \emph{makepkg}, and build a fresh OS image that is ready to flash.
The following excerpt details the three functions used to make this work.

\begin{lstlisting}[style=custombash,firstnumber=50,title=Excerpt from \emph{/scripts/build.sh}]
runMake()
{
  mkdir ./moonlight/build # Create build directory
  cd ./moonlight/build # Go to build directory
  cmake ../ # Run cmake
  make # Build the project
  sudo make install # Install the project
  sudo ldconfig /usr/local/lib # Link libraries
  cd "$parent_path" # Go back to original execution location
}

@@

In order to evaluate the performance of the project, a suite of tools must be developed to test both the newly developed application and the existing solutions.
To do this, two sets of data will be collected: First, data is recorded directly from the forked version of moonlight that logs statistics such as the number of frames received, dropped, or rendered, and data that is recorded from a separate application that logs the amount of time it takes for the client machine to receive updates after it sends data to the host machine.
The first set of data is important to track the raw performance of the application, with statistics like Frames Per Second (FPS) and decoding time which are difficult to record otherwise.
These statistics provide insight into how performant the application is from a technical perspective.
The following is a structure that holds the statistics recorded over a session of connecting to a host computer.
The data recorded is discussed in Chapter \ref{Chapter6}.

@@

The second set of data is recorded by measuring the total amount of time it takes for the host to receive data from the client, and for the client to receive updates back from the host.
This is done using the following steps once the client is connected to the host:

@@

This is sufficient to change the color of the screen back and forth between easily distinguishable colors for both the client and the host to record.
With both of these datasets recorded, the next step is to analyze the data and evaluate the performance of the application.

>>>>

\chapter{Developing the Software}

\label{Chapter5}

The second half of the project is composed of software that is used to connect the client machine to a host computer.
Designing the software begins with identifying it's requirements in Section \ref{sec:SoftwareRequirements}.
Then, potential avenues for development are discussed in Section \ref{sec:PotentialAvenues} so that the software does not need to be built from scratch.
Following that, special considerations must be made while developing the software for the CM4 running on an \emph{ARM} processor in Section \ref{sec:DevelopingForARM}, and the development of the suite of tools used to record data is detailed in Section \ref{sec:DevelopingTestingAndMeasurementTools}.
\todo{Link git repo}

@@

Unfortunately, the Nvidia GameStream protocol is both proprietary and a closed source project.
Very little is officially known about the protocol, other than it uses specific video encoders built into Nvidia Graphics cards from the 600 series from 2012 and newer \cite{gamestream_userguide}.
Apart from that, everything has been researched and reverse engineered by the community.
A GameStream session is started by first connecting to a host computer which has enabled the GameStream service through Nvidia's GeForce Experience application.
Once the service is enabled, the computer will listen on port \emph{47989} for the following web requests:
\begin{itemize}
  \item /serverinfo
  \item /pair
  \item /unpair
  \item /applist
  \item /launch
  \item /resume
  \item /cancel
\end{itemize}
Once a session has been started with a \emph{/launch} request, raw data covering video, audio, and input is passed back and forth over a number of ports until the session is terminated.
An end-to-end example of the protocol's usage is as follows:

@@

\todoadvisor{Mention this earlier, maybe to the state of the art}
Moonlight is an unofficial third-party open-source client for the Nvidia GameStream protocol \cite{moonlight}.
The project is structured as a core library written in \emph{C} with a number of community built clients built on top of it for various platforms.
In essence, it turns the process described in the previous section (\ref{subsec:NVIDIAGameStream}) into a usable programming interface.
By reverse engineering the protocol originally implemented by the Nvidia Shield, Moonlight enables the development of a GameStream client for almost any operating system.
The Moonlight community has already developed fantastic clients for many platforms such as Windows, Mac, Linux, Android, iOS, Web, and even other gaming consoles.
All that is needed for the purposes of this project is to make a fork of the project that tweaks it to best suit the CM4 that drives the hardware and develop a suite of tools to test it's functionality.

@@

In order to create an out-of-the-box experience for someone using the hardware for this project, a custom operating system image is created that has all the dependencies, libraries, and applications pre-installed and pre-configured.
Manjaro provides good tools for building and managing custom images of the Manjaro operating system for \emph{ARM} architectures \cite{manjaro_arm_tools}, with the only drawback being that it must be run on an \emph{ARM} processor, and is much more difficult to run on another architecture.
To make development easier, a toolchain was built that could be run on the CM4 to build the code that is written on other computers.
All the developer needs to do is to pull the git repository\todoadvisor{Cite the git repository} of this project onto the \emph{ARM} architecture, run a set up script to install dependencies, and then run the build script to compile everything into an immediately deployable OS image.
This build script will automatically compile all the necessary code using \emph{cmake}, package all the necessary files using \emph{makepkg}, and build a fresh OS image that is ready to flash.
The following excerpt details the three functions used to make this work.

\begin{lstlisting}[style=custombash,firstnumber=50,title=Excerpt from \emph{/scripts/build.sh}]
runMake()
{
  mkdir ./moonlight/build # Create build directory
  cd ./moonlight/build # Go to build directory
  cmake ../ # Run cmake on the project directory
  make # Build the project
  sudo make install # Install the project
  sudo ldconfig /usr/local/lib # Link libraries
  cd "$parent_path" # Go back to original execution location
}

@@

\todoadvisor{Explicitly mention that this is work I did}
In order to evaluate the performance of the project, a suite of tools must be developed to test both the newly developed application and the existing solutions.
To do this, two sets of data will be collected: First, data is recorded directly from the forked version of moonlight that logs statistics such as the number of frames received, dropped, or rendered, and data that is recorded from a separate application that logs the amount of time it takes for the client machine to receive updates after it sends data to the host machine.
The first set of data is important to track the raw performance of the application, with statistics like Frames Per Second (FPS) and decoding time which are difficult to record otherwise.
These statistics provide insight into how performant the application is from a technical perspective.
The following is a structure that holds the statistics recorded over a session of connecting to a host computer.
The data recorded is discussed in Chapter \ref{Chapter6}.

@@

The second set of data is recorded by measuring the total amount of time it takes for the host to receive data from the client, and for the client to receive updates back from the host.
This is done using the following steps once the client is connected to the host:
\todoadvisor{Mention that the times need to be syncronized}

@@

This is sufficient to change the color of the screen back and forth between easily distinguishable colors for both the client and the host to record.
\todoadvisor{Mention that this is recorded by the dataloggerz}
With both of these datasets recorded, the next step is to analyze the data and evaluate the performance of the application.

\todoadvisor{Summary}
